What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-25 android-stacks.tex
Improving introduction

The pervasiveness of the World Wide Web and the astonishing increase in the
number of smartphones have stimulated the creation of a new class of
applications which extend the phones capabilities far beyond of the basic calls
and textual messages.

Such applications run over one or more platforms, among them the Android platform
leads the smartphone market, comprising around XXXX [GoogleIO2014], thanks to its
Java-based open-source model and hardware partners like Samsung, HTC, Motorola, Asus 
(besides Nexus its own device) [ref].

Such applications are intrinsically multi-threaded and the development model and used APIs 
can be  slightly or drastically different and more error prone than tradicional development [arie-icse].

Moreover, such applications have to cope with an increasing number of abnormal
computation states. Besides the known source of failures (i.e., programmming errors, noisy 
user inputs or faults in underlying middleware or hardware) such applications need to deal with abnormal states that 
 arise as a consequence of:  memory and batery restrictions [ref], faults on third party services and libraries, 
compatibility issues caused by constant platforms updates [ref], conection with devices such as GPS, camera.

Therefore, techniques for error detection and handling are not  an optional add-on but a fundamental part of any system~\cite{bruntink2006discovering}. The Android Platform uses the exception handling mechanism embeeded
 in Java to deal with such conditions. However, although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
The consequence of the exception handling code may come the other way around making the code even less robust.

@@

Untill recent years, the availability of issues reporting exceptional conditions was scarce, while the
available data concerned a few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Along the recent years, we have witnessed a rapid increase in the number of open
source projects that make their control version and issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
When available, such information can enable different kinds of post-mortem analysis, such as:
: classify .... (get from previous versions).

Such analysis focuses a system at a time, and none of them target a family of Android Application
(which share a common core) looking for similar characteristics and vulnerabilitys on the crashes reported.

@@

Nevertheless, several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}, regardless the exception type chosen. Such guidelines do not focus on 
advocating any specific type but rather propose ways to effectively use each of them.
 We compiled a list of the guidelines related to exception types in Java, 
summarized below:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

>>>>

The pervasiveness of the World Wide Web and the astonishing increase 
in the number of smartphones stimulated the creation of applications which extend the phones
 capabilities far beyond of the basic calls and textual messages.

If we look at global smartphone shipments the numbers are astoning;
the mobile industry shiped over 315 million smartphones in the last quarter of 2013.
Such increase on smartphones combined with the pervasiveness of the World Wide Web
 have stimulated the creation of applications which extend the phones capabilities 
far beyond of the basic calls and textual messages.

In this context, Android  leads the smartphone market, having over then 1 billion 
active users per month; thanks to its open-source model and hardware partners
 like Samsung, HTC, Motorola, Asus (besides Nexus its own device) [ref].

The number of apps built on top of Android increases in a daily rate <INCLUDE SOME STATISTICS>.

%If on the one hand such applications are getting mode and more popular....
%On the other hand... there is a need for realiability, as some core functionalities 
%are moving to apps....such applications impose more challenges to software reliability

%Such mobile platforms are intrinsically multi-threaded and support a development model 
% slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],

%abnormal computation states

Such applications have to cope with an increasing number of abnormal
computation states, arising from: programmming errors (e.g. null dereferences);
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries, compatibility issues [ref], 
conection with devices e.g., as GPS, camera, among others; memory and batery restrictions
among others.

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embeeded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
The consequence of the exception handling code may come the other way around making the code even less robust.

@@

Untill recent years, the availability of issues reporting abnormal
computation states (e.g., crashes) was scarce, while the
available data concerned a few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Along the recent years, we have witnessed a rapid increase in the number of open
source projects that make their control version and issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

%Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
%found as a result of application executions. Such issues usually contain exception stack traces [ref][ref].

When available, such information can enable different kinds of post-mortem analysis, such as:
: classify .... (get from previous versions). Previous studies on crash report analysis focused a system 
at a time. Moreover, none of them target 
a family of Android applications (which share a common core) extracting the characteristics on exceptional issues 
and looking for similar vulnerabilities on applications and underlaying platform.

@@

Nevertheless, several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective} regardless the exception type used. Such guidelines do not focus on 
advocating any specific type but rather propose ways to effectively use each of them.
 We compiled a list of the guidelines related to exception types in Java, 
summarized below:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}
